Investigating the outer reaches of the Sculptor and Ursa Minor dwarf galaxies

### 2025-08-13 thesis/appendix.tex
thesis updates and ludlow notebook

>>>>

\begin{figure}
\centering
\pandocbounded{\includegraphics[keepaspectratio]{/Users/daniel/thesis_2/figures/analytic_profile_comparison.pdf}}
\caption[Comparison of analytic density profiles]{A comparison of
different parameterizations for dwarf galaxy density profiles. Note that
deviations between profiles only become apparent past 3 R\_h, and only
the Plummer profile, in contrast to more commonly assumed profiles,
deviates by \textasciitilde1 dex positive before 6 Rh. Since this
profile is a far minority in the literature, deviations from exponential
and close relatives are interesting and worth further consideration.}
\end{figure}

### 2025-08-13 thesis/discussion.tex
thesis updates and ludlow notebook

A number of works have previously speculated that dwarf galaxies such as
Sculptor and Ursa Minor may have signatures of tidal perturbations,
however not without contention. While \citet{hodge1961} and
\citet{demers+krautter+kunkel1980} were some of the earliest work to
derive the density profile of Sculptor, \citet{innanen+papp1979} was
perhaps the first to speculate that Sculptor may harbour a substantial
population of ``extratidal stars'' (stars beyond the tidal radius),
finding candidate members out to 180' in an elongated distribution from
\citet{vanagt1978}`s catalogue of variable stars.\footnote{Interestingly,
  \citet{innanen+papp1979} also speculate about Uranus's satellite
  distribution, covering gravitational dynamics at two very different
  scales.} Additionally, \citet{eskridge1988} showed a possible excess
of stars in Sculptor relative to a \citet{king1962} and Exponential
density profile beginning around 50', but suggest that this excess may
not be unusual. Later work by \citet{IH1995}, \citet{walcher+2003}, and
\citet{westfall+2006} also showed evidence of an extended component of
Sculptor's density profile, interpreting these stars as either evidence
of tidal debris or a dwarf galaxy halo. On the other hand,
\citet{coleman+dacosta+bland-hawthorn2005} show that Sculptor is well
described by a two component density profile, they additionally mention
that it is unlikely many of Sculptor's stars (less than 2\%) are
extratidal. While \citet{munoz+2018} is the most recent and deepest
photometric study of Sculptor, they only cover an area of out to radii
of about 30', so they are unable to study the possible extended
components of Sculptor detected in past works. In addition to the
presence of ``extratidal stars'', many studies note that Sculptor's
stars appear to become more elliptical with radius, consistent with
tidal effects (\citet{IH1995}; \citet{westfall+2006}).

@@

\begin{itemize}
\tightlist
\item
  \citet{revaz+jablonka2018}
\item
  \citet{tau+vivas+martinez-vazquez2024} (RRL stars )
\item
  \citet{roderick+2016} Boötes I
\item
  \citet{mcconnachie+penarrubia+navarro2007} (velocity space \& multiple
  populations).
\end{itemize}

>>>>

A number of works have previously speculated that dwarf galaxies such as
Sculptor and Ursa Minor may have signatures of tidal perturbations,
however not without contention. While \citet{hodge1961} and
\citet{demers+krautter+kunkel1980} were some of the earliest work to
derive the density profile of Sculptor, \citet{innanen+papp1979} was
perhaps the first to speculate that Sculptor may harbour a substantial
population of ``extratidal stars'' (stars beyond the tidal radius),
finding candidate members out to 180' in an elongated distribution from
\citet{vanagt1978}`s catalogue of variable stars.\footnote{Interestingly,
  \citet{innanen+papp1979} also speculate about Uranus's satellite
  distribution, covering gravitational dynamics at two very different
  scales.} Additionally, \citet{eskridge1988} showed a possible excess
of stars in Sculptor relative to a \citet{king1962} and Exponential
density profile beginning around 50', but suggest that this excess may
not be unusual. Later work by \citet{IH1995}, \citet{walcher+2003}, and
\citet{westfall+2006} also showed evidence of an extended component of
Sculptor's density profile (among other dwarfs), interpreting these
stars as either evidence of tidal debris or a dwarf galaxy halo. On the
other hand, \citet{coleman+dacosta+bland-hawthorn2005} show that
Sculptor is well described by a two component density profile, they
additionally mention that it is unlikely many of Sculptor's stars (less
than 2\%) are extratidal. While \citet{munoz+2018} is the most recent
and deepest photometric study of Sculptor, they only cover an area of
out to radii of about 30', so they are unable to study the possible
extended components of Sculptor detected in past works. In addition to
the presence of ``extratidal stars'', many studies note that Sculptor's
stars appear to become more elliptical with radius, consistent with
tidal effects (\citet{IH1995}; \citet{westfall+2006}).

@@

\begin{itemize}
\tightlist
\item
  \citet{revaz+jablonka2018}
\item
  \citet{tau+vivas+martinez-vazquez2024} (RRL stars )
\item
  \citet{roderick+2016} Boötes I
\item
  \citet{mcconnachie+penarrubia+navarro2007} (velocity space \& multiple
  populations).
\item
  \citet{cicuendez+battaglia2018}
\end{itemize}
